Instant Continuation Marks--Just Add $\lambda$

### 2012-10-05 paper.tex
Small abstract change.

We present direct and CPS-like transformations from the call-by-value $\lambda$-calculus augmented with continuation marks to the pure call-by-value $\lambda$-calculus. We discuss how such transformations simplify the construction of compilers which treat continuation marks correctly. We document how Redex, a domain-specific language for exploring language semantics, aided the discovery of correct transformations. Finally, we verify the transformations with the proof of a meaning-preservation theorem.
\end{abstract}

@@

Tail-call elimination is not part of the JavaScript specification which complicates the treatment of the tail-call behavior. We have chosen to perform a simple static analysis which determines if the recursive call is in tail position. We simulate proper tail-call behavior by statically assigning the value of the flag to reflect this.

@@

A transformation of this function desugars these constructs into vanilla JavaScript, leveraging native arrays for lists of continuation marks.
\begin{verbatim}
var fac = function( flag, marks ) {
  return function( n ) {
    if( n == 0 ) {
      console.log( (function( flag, marks ) { return marks; })( false, marks ) );
      return 1;
    }
    else {
      return (function( flag, marks ) {
        return (function( mark_value, rest_marks ) {
          return n * fac( false, [ mark_value ].concat( rest_marks ) )( n - 1 );
        })( n, flag ? marks.slice( 1 ) : marks );
      })( flag, marks );
    }
  }
}
\end{verbatim}

@@

@@

Monads, introduced by Moggi \cite{moggi1989computational}, are another way to model semantics. Their appeal is significant: they allow one to implement a particular model of computation with with a pure computational logic, such as the \lc. Ager et al. \cite{ager2003functional} implemented a stack-inspection monad in terms of a lifted state monad, but the size of the continuation must be managed by hand within.

>>>>

We present direct and CPS-like transformations from the call-by-value $\lambda$-calculus augmented with continuation marks to the pure call-by-value $\lambda$-calculus. We discuss how such transformations simplify the construction of compilers which treat continuation marks correctly. We document how Redex, a domain-specific language for exploring language semantics, aided the discovery of correct transformations. We verify the transformations with the proof of a meaning-preservation theorem. Finally, we apply the transformation to JavaScript.
\end{abstract}

@@

Tail-call elimination is not part of the JavaScript specification which complicates the treatment of the tail-call behavior. We simulate proper tail-call behavior by using the flag to encode whether a function call is in tail position.

@@

A direct-style transformation of this function desugars these constructs into vanilla JavaScript, leveraging native arrays for lists of continuation marks.
\begin{verbatim}
var fac = function( flag, marks ) {
  return function( n ) {
    if( n == 0 ) {
      console.log( (function( flag, marks ) { return marks; })( false, marks ) );
      return 1;
    }
    else {
      return (function( flag, marks ) {
        return (function( mark_value, rest_marks ) {
          return n * fac( false, [ mark_value ].concat( rest_marks ) )( n - 1 );
        })( n, flag ? marks.slice( 1 ) : marks );
      })( flag, marks );
    }
  }
}
\end{verbatim}

@@

In this example, the API to compute factorials has changed, but the change is localized. With top-level control of a program, we can transform the relevant part of the system, leaving the rest the same. This preserves interoperability with foreign functions, e.g., third-party JavaScript libraries.

@@

Monads, introduced by Moggi \cite{moggi1989computational}, are another way to model semantics. Their appeal is significant: they allow one to implement a particular model of computation with with a pure computational logic, such as the \lc.

Expressing continuation marks in a monadic way is simple enough. In Haskell,
\begin{verbatim}
data CM m a = CM ((Bool,[m]) -> a)

instance Monad (CM m) where
  return x = CM (\_ -> x)
  (CM m) >>= f = CM (\(flag,vs) -> let (CM m') = f (m (False,vs)) in m' (flag,vs))

wcm :: m -> CM m a -> CM m a
wcm v (CM m) = CM (\(f,vs) -> m (True, v:(if f then (tail vs) else vs)))

ccm :: CM m [m]
ccm = CM (\(_,vs) -> vs)

runCM :: CM m a -> a
runCM (CM m) = m (False,[])
\end{verbatim}
is one way, using our approach. The factorial functions leveraging this implementation exhibit the correct continuation mark behavior but this is \emph{not} a monad. Specifically, the right identity law
\begin{verbatim}
m >>= return = m
\end{verbatim}
does not hold since \texttt{CM} conceptually adds a stack frame in bind (\texttt{>>=}). Ager et al. \cite{ager2003functional} implemented a similar stack-inspection monad in terms of a lifted state monad, but resorted to manual management of the stack frames.
